Using Deep Learning to Identify Cognitive Impairment in Electronic Health Records

### 2021-09-07 Submission/Paper/ML4H/Paper.tex
added title and paper status
fixed the issue of weird spacing between appendix elements and added tf-idf feature correlation table

 % The booktabs package is used by this sample document
 % (it provides \toprule, \midrule and \bottomrule).
 % Remove the next line if you don't require it.
\usepackage{tabto}    
\usepackage{graphicx}
\graphicspath{ {images/} }

\usepackage{natbib}
\usepackage{url}

@@

 % change the arguments, as appropriate, in the following:
\jmlrvolume{LEAVE UNSET}
\jmlryear{2021}
\jmlrsubmitted{LEAVE UNSET}
\jmlrpublished{LEAVE UNSET}
\jmlrworkshop{Machine Learning for Health (ML4H) 2021} % W&CP title

 % The optional argument of \title is used in the header
\title[Short Title]{Full Title of Article\titlebreak This Title Has
A Line Break}

@@

\section{Introduction}
\label{sec:intro} The possibility of early diagnosis of dementia is important for improving clinical outcomes and patient management of dementia. Often, a diagnosis is given once patient has reached moderate dementia, but irreversible damage has already been done to the brain. 
% https://www.dementiacarecentral.com/aboutdementia/facts/stages/ %
Yet, dementia is not formally diagnosed or coded in claims data for over 50\% of older adults living with probable dementia. Tools that can efficiently and effectively analyze medical records for warning signs of dementia and recommend patients for follow up with a specialist can be critical to obtaining early diagnosis for dementia. We aim to use NLP to detect signs of cognitive impairment from unstructured clinician notes by using deep learning techniques.

@@

% Figure 1 - Example Sequences
\begin{figure*}[h!]
\label{fig:fig1}
\centering 
\includegraphics[width = 15cm, height = 5cm]{Figure-1.png}
\caption{Example Sequences}
\end{figure*}

@@

\label{sec:Transformer}  
\paragraph{(2) Transformer Based Sequence Classification Language Model} We utilized a pre-trained language model called Clinical-BERT, which was trained on the MIMIC II database containing EHR records from ICU patients. We used the implementation in the Huggingface Transformer and Simpletransformers packages. After text preprocessing, input texts were tokenized with the default tokenizer and converted to embeddings. The model was initialized with pre-trained parameters and later fine-tuned on our labeled training set. Hyperparameters were tuned on the held-out validation set for the cognitive impairment classification task.

@@

% Appendix A
\section{UI of Annotation Tool}
Placeholder

\newpage

% Appendix B
\section{Keywords} % wi
\begin{table}[hbtp] 
\floatconts
{tab:keywords}
    {\begin{tabular}{lccc}
        \toprule
        \bfseries Keyword & \bfseries Match Count\\
        \midrule

@@

\end{table}

\newpage

% Appendix C

>>>>

 % The booktabs package is used by this sample document
 % (it provides \toprule, \midrule and \bottomrule).
 % Remove the next line if you don't require it.
\usepackage{tabto}    
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage{float}
\usepackage{natbib}
\usepackage{url}

@@

 % change the arguments, as appropriate, in the following:
\jmlrvolume{ML4H Extended Abstract Arxiv Index}
\jmlryear{2021}
\jmlrsubmitted{201}
\jmlrpublished{}
\jmlrworkshop{Machine Learning for Health (ML4H) 2021} % W&CP title

 % The optional argument of \title is used in the header
\title[NLP Techniques to Detect Cognitive Impairment]{Using Deep Learning to Identify Cognitive Impairment in Electronic Health Records}

@@

\section{Introduction}
\label{sec:intro} The possibility of early diagnosis of dementia is important for improving clinical outcomes and patient management of dementia. Often, a diagnosis is given once patient has reached moderate dementia, but irreversible damage has already been done to the brain. 
% https://www.dementiacarecentral.com/aboutdementia/facts/stages/ %
Yet, dementia is not formally diagnosed or coded in claims data for over 50\% of older adults living with probable dementia. Tools that can efficiently and effectively analyze medical records for warning signs of dementia and recommend patients for follow up with a specialist can be critical to obtaining early diagnosis for dementia. We aim to use NLP to detect signs of cognitive impairment from unstructured clinician notes by using deep learning techniques. An overview of our project can be found in Appendix A.

@@

% Figure 1 - Example Sequences
\begin{figure*}[h!]
\label{fig:fig1}
\centering 
\includegraphics[scale = 0.35]{Figure-1.png}
\caption{Example Sequences}
\end{figure*}

@@

\label{sec:Transformer}  
\paragraph{(2) Transformer Based Sequence Classification Language Model} We utilized a pre-trained language model called Clinical-BERT, which was trained on the MIMIC II database containing EHR records from ICU patients. We used the implementation in the Huggingface Transformer and Simpletransformers packages. After text preprocessing, input texts were tokenized with the default tokenizer and converted to embeddings. The model was initialized with pre-trained parameters and later fine-tuned on our labeled training set. Optuna was used to perform a 20-trial study and tune the learning rate, adam epilson, and the number of train epochs on the held-out validation set.

@@

% Appendix A
% Figure 2 - Overview
\section{Overview}
\label{fig:fig2}
\begin{figure}[h]
\floatplacement{figure*}[h]
\centering 
\includegraphics[scale = 0.14]{Overview.png}
\caption{Overview}
\end{figure}

\clearpage

% Appendix B
\section{Keywords} 
\begin{table}[hbtp] 
\floatconts
{app:keywords}
    {\begin{tabular}{lccc}
        \toprule
        \bfseries Keyword & \bfseries Match Count\\
        \midrule

@@

\end{table}
\clearpage

% Appendix C
\section{Top TF-IDF Word Features}
\begin{table}[hbtp]
    {\begin{tabular}{cccc}
    \toprule 
    \bfseries Word & \bfseries Corr & \bfseries Word & \bfseries Corr &
    \midrule

    \textbf{Intact} & \fseries 0.56 & \textbf{Experiences} & \fseries 0.36 \\
    \textbf{Oriented} & \fseries 0.43 & \textbf{Associations} & \fseries 0.36 \\ 
    \textbf{Concentration} & \fseries 0.42 & \textbf{Homicidal} & \fseries 0.36 \\ 
    \textbf{Orientation} & \fseries 0.41 & \textbf{Observation} & \fseries 0.36 \\
    \textbf{Sensorium} & \fseries 0.40 & \textbf{Knowledge} & \fseries 0.36 \\ 
    \textbf{Perceptions} & \fseries 0.40 & \textbf{Abstract} & \fseries 0.36 \\
    \textbf{Judgement} & \fseries 0.39 & \textbf{Suicidal} & \fseries 0.35 \\
    \textbf{Fund} & \fseries 0.38 & \textbf{Attention} & \fseries 0.35 \\
    \textbf{Insight } & \fseries 0.36 & \textbf{Content} & \fseries 0.34 \\ 
    \textbf{Ideation} & \fseries 0.36 & \textbf{Thought} & \fseries 0.34  \\

    \bottomrule
  \end{tabular}}
\end{table}
